---
title: "Examine SVS Regions"
author: "Brian S Yandell"
date: "June 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Convert SVS CSV file to 3 RDS files.

```{r message=FALSE}
rawpath <- "."
library(dplyr)
```

### Read in SVS info from 28strains GZ file

Source: ftp://ftp-mouse.sanger.ac.uk/current_svs/28strains.REL-1410-SV.sdp.tab.gz

```{r}
chr <- c(1:19, "X")
cc_founders <- c("A/J", "C57BL/6J", "129S1/SvImJ", "NOD/ShiLtJ", "NZO/HlLtJ",
                 "CAST/EiJ", "PWK/PhJ", "WSB/EiJ")
strains <- sub("/", "_", cc_founders[-2])
n_strains <- length(strains)
file <- file.path(rawpath, "28strains.REL-1410-SV.sdp.tab.gz")

svs.head <- as.character(unlist(read.table(file, skip=6, comment="", nrows=1)))
svs.head[1] <- substring(svs.head[1],2)
svs.head[1:4] <- tolower(svs.head[1:4])
svs <- read.table(file)
names(svs) <- svs.head

## Get CC Founders
ccf <- c("A_J", "C57BL_6NJ", "129S1_SvImJ", "NOD_ShiLtJ", "NZO_HlLtJ", "CAST_EiJ", "PWK_PhJ", "WSB_EiJ")
svs <- cbind(svs[,1:4], svs[,ccf])
names(svs)[-(1:4)] <- cc_founders
```

### Routines to reorganize raw SVS data

```{r}
svs_length <- function(s) {
  out <- rep(0, length(s))
  indel <- (s != 0)
  s <- strsplit(s[indel], ":|-|;") ## 1:chr, 2:3:bp, 4:type
  bp <- sapply(s, function(x) 1+diff(as.numeric(x[2:3])))
  out[indel] <- unlist(bp)
  out
}
svs_length_start <- function(s) {
  out <- rep(0, length(s))
  indel <- (s != 0)
  s <- strsplit(s[indel], ":|-|;") ## 1:chr, 2:3:bp, 4:type
  bp <- sapply(s, function(x) paste(1+diff(as.numeric(x[2:3])),
                                    x[2], sep="_"))
  out[indel] <- unlist(bp)
  out
}
svs_start <- function(s) {
  out <- rep(0, length(s))
  indel <- (s != 0)
  s <- strsplit(s[indel], ":|-|;") ## 1:chr, 2:3:bp, 4:type
  bp <- sapply(s, function(x) as.numeric(x[2]))
  out[indel] <- unlist(bp)
  out
}
svs_end <- function(s) {
  out <- rep(0, length(s))
  indel <- (s != 0)
  s <- strsplit(s[indel], ":|-|;") ## 1:chr, 2:3:bp, 4:type
  bp <- sapply(s, function(x) as.numeric(x[3]))
  out[indel] <- unlist(bp)
  out
}
svs_type <- function(s) {
  indel <- (s != 0)
  s <- strsplit(s[indel], ":|-|;") ## 1:chr, 2:3:bp, 4:type
  paste(unique(sapply(s, function(x) x[4])), collapse=",")
}
```

### Condense full data

#### First find start and end for 8 CCFs for each SVS

Reduce down to SVS with something going on.

```{r}
svs_cc_type <- apply(svs[,-(1:4)], 1, svs_type)
svs <- svs[svs_cc_type != "",]
svs_cc_type <- svs_cc_type[svs_cc_type != ""]
```

Identify start and end for each CCF and SVS

```{r}
svs_cc_start <- apply(svs[,-(1:4)], 2, svs_start)
svs_cc_end <- apply(svs[,-(1:4)], 2, svs_end)
```

Some really big intervals

```{r}
tmp <- apply(svs_cc_start, 1, function(x) min(x[x>0]))
summary(tmp-svs$start)
```

```{r}
svs[which((tmp-svs$start)>30000),]
apply(svs[which((tmp-svs$start)>30000),2:3],1,diff)
```

Construct desired data frame.

```{r}
svs8 <- svs[,1:3]
svs8$start <- apply(svs_cc_start, 1, function(x) min(x[x>0]))
svs8$end <- apply(svs_cc_end, 1, max)
svs8$type <- svs_cc_type
```

Some `end` values are beyond `B6` interval. Look at them.
It seems to make sense to use the CCF ends

```{r}
data.frame(svs[,1:3],svs_cc_end)[which((svs$end-svs8$end) < 0),]
```

Arrange SVS types by maximum length.

```{r}
svs8 %>%
  group_by(type) %>%
  summarize(count=n(),
            min_len=min(end-start),
            max_len=max(end-start)) %>%
  arrange(desc(max_len))
```

Arrange SVS types by number of each type.

```{r}
svs8 %>%
  group_by(type) %>%
  summarize(count=n(),
            min_len=min(end-start),
            max_len=max(end-start)) %>%
  arrange(desc(count))
```

Now make three data frames to save

```{r}
svs8_start <- cbind(svs8, svs_cc_start)
svs8_end <- cbind(svs8, svs_cc_end)
svs_cc_diff <- svs_cc_end - svs_cc_start + 1
svs_cc_diff[svs_cc_diff == 1] <- 0
svs8_len <- cbind(svs8, svs_cc_diff)
saveRDS(svs8_start,file=file.path(datapath,"svs8_start.rds"))
saveRDS(svs8_end,file=file.path(datapath,"svs8_end.rds"))
saveRDS(svs8_len,file=file.path(datapath,"svs8_len.rds"))
```
